---
title: "Options Pricing using Binomial Trees in R"
author: "Bashir Taiwo"
date: "10/12/2020"
output:
  word_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```



\
\

This report demonstrates how to calculate American and European option prices using R. Two different examples were considered, and the report was generated using R Markdown\
\

                      
## Example (European Call Option)\                     
A stock price is currently $50. Over each of the next two 3-month periods it is expected
to go up by 6% or down by 5%. The risk-free interest rate is 5% per annum with
continuous compounding. What is the value of a 6-month European call option with a
strike price of $51?\

Stock price (s) = $50\
Time (tt) = 3/12\
rate (r) = 5%\
upward movement (up) = 1.06\
downward movement (dn) = 0.95\
Strike price (k) = $51\

Now, using binomopt function in R, we can calculate the price of the European call option

```{r}
library(derivmkts)
binomopt(s=50,k=51, up=1.06,dn=0.95, r=0.05, tt=0.5, d=0, nstep=2, american=F, putopt=F,crr=F,specifyupdn=T)
```

The call option price is $1.635071. 

Binomial trees were then plotted using binomplot function in the code below:

```{r}

binomplot(s=50,k=51,v=0,up=1.06,dn=0.95, r=0.05, tt=0.5, d=0, nstep=2, putopt=F, american=F,
    plotvalues=T, plotarrows=T, drawstrike=TRUE,
    pointsize=4, ylimval=c(0,0),
    saveplot = FALSE, saveplotfn='binomialplot.pdf',
    crr=FALSE, jarrowrudd=FALSE, titles=TRUE, specifyupdn=T, returnprice=FALSE, logy=FALSE)

```


In the case of an American put option, the following question was be used to illustrate how to calculate American put option in R

\
## Example (American Put Option)\
Suppose  the  current  stock  price  is  $100  with  a  strike  price  of  $110  and  the expiration  date  to  be  2  years.   The  risk  free  rate  is  5%  per  annum  with  continuous compounding.  The stock prices increases by 8% and decreases by 6%.  Use a 4-step tree to value an American put.\


Stock price (s) = $100\
Time (tt) = 24/12\
rate (r) = 5%\
upward movement (up) = 1.08\
downward movement (dn) = 0.94\
Strike price (k) = $110\
```{r}
binomopt(s=100,k=110, up=1.08,dn=0.94, r=0.05, tt=2, d=0, nstep=4, american=T, putopt=T,crr=T,specifyupdn=T)
```

The put option price is $10. 

Below is a plot of the corresponding binomial trees

```{r}
binomplot(s=50, k=51,r=0.05, tt=2, v=0, d=0, nstep=4, putopt=T, american=TRUE,
    plotvalues=T, plotarrows=T, drawstrike=TRUE,
    pointsize=4, ylimval=c(0,0),
    saveplot = FALSE, saveplotfn='binomialplot.pdf',
    crr=T, jarrowrudd=FALSE, titles=TRUE, specifyupdn=T,
    up=1.06, dn=0.95, returnprice=FALSE, logy=FALSE)

```


